---
title: "Course evaluations and simple linear regression"
author: "Maxwell Waun"
date: "March 30, 2020"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

<style type="text/css">
/* Title */
h1.title {
  color: #08738a;
  font-size:60px;
  font-weight: bold;
}
/* Level 1 header */
h1 {
  color: #0ba2c2;
}
/* Level 2 header */
h2 {
  color: #0ba2c2;
}
/* Level 4 header */
h4 {
  font-weight: bold;
}
/* Table of contents */
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    z-index: 2;
    color: #fff;
    background-color: #82D0C7;
    border-color: #337ab7;
}
    
</style>

```{r setup, include=FALSE}
# global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      comment = NA,
                      fig.width=6, 
                      fig.height=5)
```

# Introduction

This is adapted from Lab 6 in Duke's Introduction to Data Science course.

We will analyze what goes into course evaluations and how certain variables
effect the overall score.

To get started, load packages `tidyverse` and `broom`. Install
any packages with code `install.packages("package_name")`.

```{r packages, echo=TRUE}
library(tidyverse)
library(broom)
```


```{r theme-custom, include=FALSE}
theme_custom <- function() {
  theme_bw() +
  theme(text = element_text(color = "#4C6A92"),
        axis.title = element_text(size = 16), 
        title = element_text(size = 20),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.caption = element_text(size = 10))
}
```

# Data

The data were gathered from end of semester student evaluations for a 
large sample of professors from the University of Texas at Austin. 
In addition, six students rated the professors' physical appearance. 
Each row in `evals` contains a different course and the 
columns represent variables about the courses and professors.

Use `read_csv()` to read in the data and save it as an object named `evals`.
The data is available on D2L.


```{r data, echo=TRUE}
evals <- read_csv("evals-mod.csv")
```

## Data dictionary

**Variable** | **Description**
-------------|---------------------------------------------------------
score |	Average professor evaluation score: (1) very unsatisfactory - (5) excellent
rank |	Rank of professor: teaching, tenure track, tenure
ethnicity |	Ethnicity of professor: not minority, minority
gender |	Gender of professor: female, male
language |	Language of school where professor received education: english or non-english
age |	Age of professor
cls_perc_eval |	Percent of students in class who completed evaluation
cls_did_eval |	Number of students in class who completed evaluation
cls_students |	Total number of students in class
cls_level |	Class level: lower, upper
cls_profs |	Number of professors teaching sections in course in sample: single, multiple
cls_credits	| Number of credits of class: one credit (lab, PE, etc.), multi credit
bty_f1lower |	Beauty rating of professor from lower level female: (1) lowest - (10) highest
bty_f1upper |	Beauty rating of professor from upper level female: (1) lowest - (10) highest
bty_f2upper |	Beauty rating of professor from upper level female: (1) lowest - (10) highest
bty_m1lower |	Beauty rating of professor from lower level male: (1) lowest - (10) highest
bty_m1upper |	Beauty rating of professor from upper level male: (1) lowest - (10) highest
bty_m2upper |	Beauty rating of professor from upper level male: (1) lowest - (10) highest

# Exploratory data analysis

#### Task 1

Create a scatter plot between `score` and a quantitative variable. Comment
on the relationship. What do you think the correlation is between the variables?
```{r task1}
ggplot(evals, aes(x = score, y = cls_students)) + geom_point() + theme_custom()
#There isn't a clear relationship, but it looks as though the more students the professor teaches, the better likelihood he/she will get a higher score.
```
#There isn't a clear relationship, but it looks as though the more students the professor teaches, the better likelihood he/she will get a higher score.


#### Task 2

Create a scatter plot between `score` and a categorical variable. Comment
on the relationship. Add `geom_jitter()` to your plot.
What do you notice about the variability in score
between different levels?
```{r task2}
ggplot(evals, aes(x=score, y=cls_level)) + geom_point() + geom_jitter() + theme_custom()
#it appears that teaching both upper and lower level students gives the same looking graph, suggesting there is little difference made.
```
#it appears that teaching both upper and lower level students gives the same looking graph, suggesting there is little difference made.


#### Task 3

Create a variable named `avg_bty`. This is the average attractiveness from the
six student scores for each professor. Add this variable to `evals`.
*Hint*: `rowwise()` then `mutate()` then `ungroup()`
```{r task3}
evals = evals %>% 
  rowwise() %>%
  mutate(avg_bty = mean(c(bty_f1lower,bty_f1upper,bty_f2upper,bty_m1lower,bty_m1upper, bty_m2upper))) %>%
  ungroup()

```


#### Task 4

Create a scatter plot between `score` and `avg_bty`. Comment
on the relationship. What do you think the correlation is between the variables?
```{r task4}
ggplot(evals, aes(x=score, y=avg_bty)) + geom_point() + theme_custom()
#There seems to be a slight correlation between the variables. The relationship looks linearly positive, suggesting that those who were seen as better looking reveived slightly better scores. 
```
#There seems to be a slight correlation between the variables. The relationship looks linearly positive, suggesting that those who were seen as better looking reveived slightly better scores. 


#### Task 5

Modify your plot from Task 4. Add a layer with `geom_jitter()`. Comment
on the relationship.
```{r task5}
ggplot(evals, aes(x=score, y=avg_bty)) + geom_point() + geom_jitter() + theme_custom()
#Its easier to view in this format. The relationship looks the same as before, that professers seen as better looking tend to get better scores on their evaluations.
```
Its easier to view in this format. The relationship looks the same as before, that professers seen as better looking tend to get better scores on their evaluations.


# Model build

#### Task 6

Fit a linear model with function `lm()` between variables `score` and
`avg_bty`. Save it as an object `m.bty`.
What are the estimates for $b_0$ and $b_1$? *Hint*: `tidy()`
```{r task6}
m.bty <- lm(score ~ avg_bty, data = evals)
m.bty

```
b_0 = 3.88033
b_1 = 0.06664


#### Task 7

Modify your plot from Task 4 and include the regression line. Pick a different
color than the default of blue.
```{r task7}
evals %>%
  ggplot(mapping = aes(x = score, y = avg_bty)) + 
  geom_point() + stat_smooth(method = "lm", se = FALSE, color = "red") +
  theme_custom()
```


#### Task 8

Interpret the slope of the fitted model. Does the intercept have any practical
meaning within the scope of our data?

Yes the intercept (if I'm understanding the question right) represents the avg_bty score so it does have practial meaning. If the regression line was placed higher or lower it would represent a different thing. 

#### Task 9

What is the $R^2$ value? Interpret this value. *Hint*: `glance()`
```{r task9}
m.bty %>%
  broom::glance()

```
R^2 = 0.03402322
This isn't good. This means that we have a lot of deviation from our line as you can see from the graph.


# Residuals

#### Task 10

Create a plot of the residual value versus `avg_bty`. Comment on what you
observe.
```{r}
resid <- m.bty %>% broom::augment()
resid <- resid$.resid
ggplot(evals, aes(x = avg_bty, y = resid)) + geom_point() + theme_custom()
#I don't know what conclusions to make of this
```
I don't know what conclusions to make of this


#### Task 11

Compute the sum of squared residuals. Function `augment()` will give you all
the residuals.
```{r}
sum(resid**2)
```



# References {#References-link}

1. http://www2.stat.duke.edu/courses/Spring18/Sta199/labs/lab-06-modelling-course-evals.html

